Real-Time Application of Activity Classification from Body-Worn Accelerometers

### 2016-06-06 notes/paper/paper.tex
Added more detail to the paper. More specifically, more mathematical detail.

\documentclass[journal]{IEEEtran}
\usepackage{blindtext}
\usepackage{graphicx}
\usepackage[english]{babel}
\usepackage{pinyin}
\usepackage{natbib}
\usepackage{notoccite}
\usepackage{amssymb}
\usepackage{textcomp}
\usepackage{floatrow}
%
\begin{document}
%
% paper title
\title{Real-Time Application of Activity Classification from Body-Worn Accelerometers}
%
\author{Craig~Euler, C.T. Lin, Melissa Flores, et al? (Ruben, Brian, Sydney, Steve?)}
%
% make the title area
\maketitle
%
\begin{abstract}
Body worn sensors have been used in the market for many years now.
Many of which are used for activity monitoring or fitness.
In this paper, we make use of a matched-filtering algorithm for real-time activity classification from data obtained with three-axis accelerometers worn by the user.
We achieve invariance to sensor orientation with dimensionality reduction.
We make use of an instance-based learning algorithm to train the device to learn the individual's motion patterns and store that information as activity filters for our matched filter.
We improve computational efficiency with dimensionality reduction, temporal decomposition, and pre-processing of our filters for later use.
An assessment of our computational costs are provided as processing flop counts for the real-time specific parts of the algorithms.
\end{abstract}
%
\section{Introduction}
Various methods for activity classification exist.
Custom dicision tree, automatically generated decision tree, and artificial neural networks \cite{parkka_ermes_korpipaa_mantyjarvi_peltola_korhonen_2006} have been explored as well as other frequency based classification methods \cite{sharma_purwar_lee_lee_chung_2008}.
In this paper, we employ the use of a matched-filter method to identify an individual’s activity in real-time.
The matched-filtering method with use of thresholding and comparing the correlations of various filters is a simple method to use and have been studied in the past \cite{giannakis_tsatsanis_1990}.
Computational resources are limited on micro-processors so we've designed our algorithms to be as efficient as possible.
Having an a priori knowledge of the activity the individual is performing is useful for many current fitness algorithms.
These devices (Fit-Bit, Apple watch, Moov, Jawbone, etc.) usually have the individual input the activity they are performing.
%
\section{Activity Classification via Matched Filtering}
The matched-filter method is commonly used for extracting a signal out of noisy data.
We employ this method for identifying a user’s activity from data generated by body-worn accelerometers.
Because the matched-filter method requires use of an existing reference template to identify the desired signal, we must develop our template from a training set.
We will discuss how we develop our set of templates for use in our matched-filter.
%
\section{Dimensionality Reduction}
We manage to reduce our computational resources and remove our dependence on sensor orientation at the same time through a process of dimensionality reduction.
We assume that over a short period of time a user’s limb movement is constrained mostly to a two-dimensional plane.
By identifying the plane of motion, we then project the acceleration data onto this plane and re-define our $x$ and $y$ axis accordingly.
By this assumption, activities that are mostly confined to two dimensions will improve our signal identification.

>>>>

\documentclass[journal]{IEEEtran}
\usepackage{blindtext}
\usepackage{graphicx}
\usepackage[english]{babel}
\usepackage{pinyin}
\usepackage{natbib}
\usepackage{notoccite}
\usepackage{amssymb}
\usepackage{textcomp}
\usepackage{floatrow}
\usepackage{amsmath}
%
\begin{document}
%
% paper title
\title{Real-Time Application of Activity Classification from Body-Worn Accelerometers}
%
\author{Craig~Euler, C.T. Lin, Melissa Flores, et al? (Ruben, Bryan, Sydney, Steve?)}
%
% make the title area
\maketitle
%
\begin{abstract}
Body worn sensors have been used in the market for many years now.
Many of which are used for activity monitoring or fitness.
In this paper, we make use of a matched-filtering algorithm for real-time activity classification from data obtained with three-axis accelerometers worn by the user.
We achieve invariance to sensor orientation with dimensionality reduction.
We make use of an instance-based learning algorithm to train the device to learn the individual's motion patterns and store that information as activity filters for our matched filter.
We improve computational efficiency with dimensionality reduction, temporal decomposition, and pre-processing of our filters for later use.
An assessment of our computational costs are provided as processing flop counts for the real-time specific parts of the algorithms.
\end{abstract}
%
\section{Introduction}
Various methods for activity classification exist.
Custom dicision tree, automatically generated decision tree, and artificial neural networks \cite{parkka_ermes_korpipaa_mantyjarvi_peltola_korhonen_2006} have been explored as well as other frequency based classification methods \cite{sharma_purwar_lee_lee_chung_2008}.
In this paper, we employ the use of a matched-filter method to identify an individual’s activity in real-time.
The matched-filtering method with use of thresholding and comparing the correlations of various filters is a simple method to use and have been studied in the past \cite{giannakis_tsatsanis_1990}.
Computational resources are limited on micro-processors so we've designed our algorithms to be as efficient as possible.
Having an a priori knowledge of the activity the individual is performing is useful for many current fitness algorithms.
These devices (Fit-Bit, Apple watch, Moov, Jawbone, etc.) usually have the individual input the activity they are performing.
%
\section{Activity Classification via Matched Filtering}
The matched-filter method is commonly used for extracting a signal out of noisy data.
We employ this method for identifying a user’s activity from data generated by body-worn accelerometers.
Because the matched-filter method requires use of an existing reference template to identify the desired signal, we must develop our template from a training set.
We will discuss how we develop our set of templates for use in our matched-filter.

The matched filter is based on the cross correlation of the data $\vec{y} = \{y\}_{k=0}^{N-1}$ with the reference signal $\vec{x} = \{x\}_{k=0}^{M-1}$.
% put this later when we refer to the convolution theorem
If we define $\vec{x}'$ to be $\vec{x}$ that has been zero padded to the length of $\vec{y}$ with the added restriction that the length of $\vec{x}$ be no more than half the length of $\vec{y}$ for preventing corruption with circular indexing, then the cross correlation of $\vec{x}$ with $\vec{y}$ is defined as:
%
\begin{equation} \label{cross_correlation_eq}
(\vec{x} \star \vec{y}) = \{\sum_{i=0}^{M}(x_{i} y_{(i+k)\%N})\}_{k=0}^{N-M}
\end{equation}
%
which is simply a moving dot product. Here we use $(i+k)\%N$ to represent $i+k$ modulo $N$.
We can then normalize \eqref{cross_correlation_eq} by the following:
%
\begin{equation} \label{norm_cross_correlation_eq}
\widehat{(\vec{x} \star \vec{y})}_k = \frac{(\vec{x} \star \vec{y})_k}{\sqrt{||\vec{x}|| \sum_{p=0}^M y_{p+k}^2}}
\end{equation}
%
where \eqref{norm_cross_correlation_eq} is now bound by $[-1,1]$ and has removed any biased toward high energy.
We are most interested where our reference best matches our data. The resulting max correlation yeilds our matched-filter output:
%
\begin{equation} \label{matched_filter_eq}
MF := \underset{k \in [0, M-N]}{max} \{\widehat{(\vec{x} \star \vec{y})}_k\}
\end{equation}
%
\section{Dimensionality Reduction}
We manage to reduce our computational resources and remove our dependence on sensor orientation at the same time through a process of dimensionality reduction.
We assume that over a short period of time a user’s limb movement is constrained mostly to a two-dimensional plane.
By identifying the plane of motion, we then project the acceleration data onto this plane and re-define our $x$ and $y$ axis accordingly.
By this assumption, activities that are mostly confined to two dimensions will improve our signal identification.
